Stability Analysis of Recurrent Neural Networks \\ using Ordinary Differential Equations

### 2020-09-22 p5/stability_rnn.tex
WIP 22Sep2020

\bibliographystyle{plain}
\bibliography{stability_rnn}
\end{document}

>>>>

\section{GRU's as Dynamical Systems}\label{s11}
The most distinctive characteristic of a recurrent neural network is that it
allows previous outputs to be used as inputs. Mathematically, one can express
them as
\begin{equation}\label{s11e1}
h_{t+1} = f_1(h_t, x_t),
\end{equation}
where $x_t$ is the input and $h_t$ is the hidden state at `time' $t$. The
function $f$ is usually non-linear. Equation \eqref{s11e1} is the general form
of a non-linear difference equation. In the absence of any input, the hidden
state evolves as
\begin{equation}\label{s11e2}
h_{t+1} = f_1(h_t, 0) = f(h_t).
\end{equation}
This is an example of an autonomous dynamical system. The evolution of the
hidden states of recurrent neural networks can thus be considered to be
trajectories of dynamical systems. One can then use the rich theory of 
dynamical systems to study the stability of recurrent neural networks.

A gated recurrent unit with the structure
\begin{eqnarray}
z_t &=& \sigma(W_zx_t + U_zh_{t-1} + b_z) \label{s11e3} \\
r_t &=& \sigma(W_rx_t + U_rh_{t-1} + b_r) \label{s11e4} \\
h_t &=& (1 - z_t)\circ\tanh(W_hx_t + U_h(r_t\circ h_{t-1} + b_h)) + 
        z_t \circ h_{t-1} \label{s11e5}
\end{eqnarray}
where $W_z, W_r, W_h \in \mathbb{R}^{d \times p}$, $U_z, U_r, U_h \in 
\mathbb{R}^{d \times d}$ are parameter matrices and  $b_z, b_r, b_h \in
\mathbb{R}^d$ are bias vectors. In the absence of an input the above
equations become
\begin{eqnarray}
z_t &=& \sigma(U_zh_{t-1} + b_z) \label{s11e6} \\
r_t &=& \sigma(U_rh_{t-1} + b_r) \label{s11e7} \\
h_t &=& (1 - z_t)\circ\tanh(U_h(r_t\circ h_{t-1} + b_h)) + 
        z_t \circ h_{t-1} \label{s11e8}
\end{eqnarray}
One can consider equations \eqref{s11e6} to \eqref{s11e8} to be discrete
forms of the system of ODEs \cite{jordan2019gated}
\begin{eqnarray}
z(t) &=& \sigma(U_zh(t) + b_z) \label{s11e9} \\
r(t) &=& \sigma(U_rh(t) + b_r) \label{s11e10} \\
\dot{h}(t) &=& (1 - z(t))\circ
	\left(-h(t)+\tanh(U_h(r(t)\circ h(t)) + b_h)\right) \label{s11e11},
\end{eqnarray}
where the dot over $h$ denotes its time derivative. We consider a special case
of these equations with the following values of parameters, (first row of 
table I in \cite{jordan2019gated},
\begin{eqnarray}
U_r &=&
\bibliographystyle{plain}
\bibliography{stability_rnn}
\end{document}
